---
title: "My Course Project"
author: "Hakan UYSAL"
date: "12/23/2019"
output: html_document
runtime: shiny
---
## Read and group
This part is to read data and put it into data frame.

Related activity data is received from below link.

[data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)



Then unzipped and put to local folder.


```{r setup, include=TRUE,eval=TRUE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/ENKHUY/Documents/TRAINING/R/Reproducible")
a<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"



library(tidyr)
library(dplyr)
library(reshape2)
library(data.table)
library(tidyselect)
library(datasets)
library(lattice)
library(ggplot2)
library(shiny)

if(!file.exists("./data/activity.csv")) 
  { 
    download.file(a,destfile = "./data/activity.zip")
    unzip("./data/activity.zip") 
  } 
data<-read.csv("./data/activity.csv")
data$date<-as.Date(data$date)
data$steps<-as.integer(data$steps)
data$interval<-as.integer(data$interval)
summary(data)
head(data,20)
```


## Calculate and Plot daily total steps
This part is to calculate and plot the sum for daily steps.


```{r}
knitr::opts_chunk$set(echo = TRUE)
data2 <- data  %>% group_by(date) %>% summarise(sum(steps))
names(data2)[2]<-"steps"
names(data2)[1]<-"date"
data2<-na.omit(data2,c("steps"))


  ggplot(data2, aes(x = date, y = steps)) +
        geom_bar(stat = "identity", fill = "blue") +
        ggtitle("Daily Total Steps") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        xlab("Date") +
        ylab("Steps")

print("Mean and Median")
print(mean(data2$steps))
print(median(data2$steps))
```
## Average daily activity pattern
In this section daily activity will be plotted.

```{r}
knitr::opts_chunk$set(echo = TRUE)

data2<-mutate(data,dateinterval=as.POSIXct(as.POSIXct(data$date)+data$interval*60))

data2 %>% 
  ggplot( aes(x=dateinterval, y=steps)) +
    geom_line(color="green") +
    ylim(0,800)   

d<-order(data2$steps,decreasing = TRUE)
# find maximum steps
print("Maximum Steps Date")
print(data2$dateinterval[d[1]])

```

## Imputing missing values

This section the missing values will be imputed.
### Mean is selected as strategy to impute values.

```{r}
knitr::opts_chunk$set(echo = TRUE)

data3<-data
nadata<-is.na(data$steps)
data3$steps[nadata]<-mean(data$steps,na.rm=TRUE)

data4 <- data3  %>% group_by(date) %>% summarise(sum(steps))
names(data4)[2]<-"steps"
names(data4)[1]<-"date"
ggplot(data4, aes(x = date, y = steps)) +
        geom_bar(stat = "identity", fill = "blue") +
        ggtitle("Daily Total Steps") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        xlab("Date") +
        ylab("Steps")
```

### Observation

It is observed that after imputation gaps on the graphs are filled.

## Are there differences between weekdays and weekends?

Weekdays and weekends data will be put to a factor variable and analysis will be done to draw both.

```{r}
knitr::opts_chunk$set(echo = TRUE)

data3<-data
nadata<-is.na(data$steps)
data3$steps[nadata]<-mean(data$steps,na.rm=TRUE)

data2<-mutate(data3,weekday=weekdays(date))
data2<-mutate(data2,weekend=(weekday=="Saturday" | weekday=="Sunday"))
data2<-mutate(data2,weekendd="weekday")
data2$weekendd[data2$weekend]<-"weekend"

# Categorise the weekend for activity days.

data2 %>% 
  ggplot( aes(x=interval,y=steps)) +
    geom_line(color="blue",size=0.2)+ 
  facet_wrap(.~weekendd,strip.position = "top",dir="v")+
  labs(title="weekend-weekday activity",x="5 minutes Interval",y="number of steps")
     
```

### Observation

Weekends actiivity is a bit less compared to weekdays activity.

